---
title: "Hands-on Exercise 01a: Geospatial Data Wrangling"
format: html
execute:
  freeze: true
  warning: false
  message: true
---

## Overview

This exercise reproduces the full workflow from Sections 1.5–1.9 of *Geospatial Data Science with R*. It walks through simple feature inspection, quick exploratory plotting, coordinate system alignment, converting the Airbnb listings table to an `sf` object, and two geoprocessing case studies using cycling path buffers and preschool distribution.

## Setup

```{r setup}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(sf, dplyr, readr, tmap, ggplot2, janitor, stringr, units, here)

path_geo <- here::here("Hands-on_Ex", "Hands-on_ex01", "data", "geospatial")
path_asp <- here::here("Hands-on_Ex", "Hands-on_ex01", "data", "aspatial")
```

## Step 1: Import Geospatial and Aspatial Data

```{r import-data}
mpsz <- st_read(dsn = path_geo, layer = "MP14_SUBZONE_WEB_PL")
cyclingpath <- st_read(dsn = file.path(path_geo, "CyclingPath_Aug2025"), layer = "CyclingPathGazette")
preschool <- st_read(file.path(path_geo, "PreSchoolsLocation.kml"))
listings <- read_csv(file.path(path_asp, "listings.csv"), show_col_types = FALSE)
```

## Step 2: Inspect simple feature data frames (Ch. 1.5)

```{r inspect-geometry}
st_geometry(mpsz)
```

```{r inspect-glimpse}
glimpse(mpsz)
```

```{r inspect-head}
head(mpsz, n = 5)
```

## Step 3: Produce quick exploratory plots (Ch. 1.6)

```{r plot-mpsz}
plot(mpsz)
```

```{r plot-geometry-only}
plot(st_geometry(mpsz))
```

```{r plot-attribute}
plot(mpsz["PLN_AREA_N"])
```

```{r plot-overlay-before-transform}
plot(st_geometry(mpsz))
plot(st_geometry(preschool), add = TRUE)
```

## Step 4: Align coordinate reference systems (Ch. 1.7)

```{r assign-crs}
st_crs(mpsz)
mpsz <- st_set_crs(mpsz, 3414)
st_crs(mpsz)
```

```{r transform-preschool}
st_geometry(preschool)
preschool <- st_transform(preschool, crs = 3414)
st_geometry(preschool)
```

```{r plot-overlay-after-transform}
plot(st_geometry(mpsz))
plot(st_geometry(preschool), add = TRUE, col = "red", pch = 20)
```

## Step 5: Import and convert aspatial data (Ch. 1.8)

```{r inspect-listings}
list(listings)
```

```{r listings-to-sf}
listings_sf <- listings %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(crs = 3414)

glimpse(listings_sf)
```

```{r plot-listings}
plot(st_geometry(mpsz))
plot(st_geometry(listings_sf), add = TRUE, col = "darkorange", pch = 20)
```

## Step 6: Geoprocessing case 1—cycling path buffers (Ch. 1.9.1)

```{r cycling-buffer}
buffer_cycling <- st_buffer(cyclingpath, dist = 5, nQuadSegs = 30)
buffer_cycling$AREA <- st_area(buffer_cycling)
sum(buffer_cycling$AREA)
```

```{r buffer-plot}
mpsz_selected <- mpsz %>% filter(SUBZONE_N == "TAMPINES WEST")
buffer_cycling_selected <- st_intersection(buffer_cycling, mpsz_selected)

plot(st_geometry(mpsz_selected))
plot(st_geometry(buffer_cycling_selected), col = "lightblue", add = TRUE)
```

## Step 7: Geoprocessing case 2—preschool counts and density (Ch. 1.9.2)

```{r preschool-count}
mpsz$`PreSch Count` <- lengths(st_intersects(mpsz, preschool))
summary(mpsz$`PreSch Count`)
```

```{r preschool-top}
top_n(mpsz, 1, `PreSch Count`)
```

```{r preschool-density}
mpsz$Area <- st_area(mpsz)
mpsz <- mpsz %>% mutate(`PreSch Density` = `PreSch Count` / Area * 1000000)
```

```{r preschool-hist-base}
hist(mpsz$`PreSch Density`)
```

```{r preschool-hist-ggplot}
ggplot(data = mpsz, aes(x = as.numeric(`PreSch Density`))) +
  geom_histogram(bins = 20, color = "black", fill = "lightblue") +
  labs(
    title = "Are pre-school even distributed in Singapore?",
    subtitle = "There are many planning subzones with a single pre-school,\nthere are seven planning subzones with at least 30 or more pre-schools",
    x = "Pre-school density (per km sq)",
    y = "Frequency"
  )
```

```{r preschool-scatter}
ggplot(data = mpsz, aes(x = as.numeric(`PreSch Density`), y = `PreSch Count`)) +
  geom_point(color = "black", fill = "lightblue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(
    x = "Pre-school density (per km sq)",
    y = "Pre-school count"
  )
```

## Step 8: Produce cleaned outputs and export

```{r clean-and-export}
mpsz_clean <- mpsz %>% janitor::clean_names()
cycling_clean <- cyclingpath %>% janitor::clean_names()
preschool_clean <- preschool %>% janitor::clean_names()
listings_sf_clean <- listings_sf %>% janitor::clean_names()

clean_path <- file.path(path_geo, "clean_outputs")
if (!dir.exists(clean_path)) dir.create(clean_path, recursive = TRUE)

st_write(mpsz_clean, file.path(clean_path, "subzone.gpkg"), layer = "subzone", delete_dsn = TRUE)
st_write(cycling_clean, file.path(clean_path, "cycling_path.gpkg"), layer = "cycling_path", delete_dsn = TRUE)
st_write(preschool_clean, file.path(clean_path, "preschools.gpkg"), layer = "preschools", delete_dsn = TRUE)
st_write(listings_sf_clean, file.path(clean_path, "listings_sf.gpkg"), layer = "listings_sf", delete_dsn = TRUE)
```

## Step 9: Processing summary

```{r summary}
list(
  subzone_records = nrow(mpsz_clean),
  cycling_records = nrow(cycling_clean),
  preschool_records = nrow(preschool_clean),
  listings_records = nrow(listings_sf_clean),
  total_cycling_buffer_area_sqkm = set_units(sum(buffer_cycling$AREA), "km^2"),
  preschool_density_summary = summary(as.numeric(mpsz$`PreSch Density`))
)
```
